---
title: "PA1_template"
author: "Ankit"
date: "July 31, 2016"
output: 
   html_document:
     keep_md: yes
---
## Load the libraries
```{r}
library(ggplot2)
library(plyr)
library(lubridate)
```
## Loading and preprocesssing the data
```{r}
unzip("repdata%2Fdata%2Factivity.zip")
activitydata<-read.csv("activity.csv")
#head(activitydata)
```

## What is mean total number of steps taken per day?
```{r}
#sumofstepperday<-aggregate(activitydata$steps,by=list(activitydata$date),FUN=sum)
sumofstepperday<-ddply(activitydata, c("date"),summarise,
                   totalsteps=sum(steps,na.rm=TRUE)
                   )
hist(sumofstepperday$totalsteps,xlab="SumOfStepsPerDay",main="Histogram of the total number of steps taken each day")
```
```{r,echo=FALSE}
mean<-mean(sumofstepperday$totalsteps, na.rm=TRUE)
median<-median(sumofstepperday$totalsteps)
```

The mean total number of steps taken per day is `r mean`. The median number of steps taken per day is `r median`(NA's omitted).


## What is the average daily activity pattern?
```{r}
#meanofstepsperday<-aggregate(activitydata$steps,by=list(activitydata$date),FUN=mean)
stepsper5min<-ddply(activitydata, c("interval"),summarise,
                    meansteps = mean(steps,na.rm=TRUE)
                    )
plot(stepsper5min$interval,stepsper5min$meansteps,type="l",ylab = "Mean_ Steps",xlab="5-minute interval",main="Average steps for each 5-min interval")
```
```{r,echo=FALSE}
result<-stepsper5min[which(stepsper5min$meansteps==max(stepsper5min$meansteps)),"interval"]
```

The five minute interval with the highest mean step-count is interval:
`r result`


## Imputing missing values
```{r,echo=FALSE}
missing<-sum(is.na(activitydata$steps))
```
There are `r missing` missing values in the dataset.

Strategy to fill all the mssing values.

```{r}
step_interpolation <- function(rownumber){
  prevrow=rownumber;
  nextrow=rownumber;
  while(is.na(activitydata$steps[prevrow])){
    prevrow=prevrow-1
    if(prevrow<1)return(mean(activitydata[activitydata$interval==activitydata$interval[rownumber],"steps"],na.rm=TRUE))
  }
  while(is.na(activitydata$steps[nextrow])){
    nextrow=nextrow+1
    if(nextrow>nrow(activitydata))return(mean(activitydata[activitydata$interval==activitydata$interval[rownumber],"steps"],na.rm=TRUE))
  }
  return(
    (activitydata$steps[prevrow]+activitydata$steps[nextrow])/2
  )
}

activity_guessNA <-activitydata
for(n in 1:nrow(activitydata)){
  if(is.na(activitydata$steps[n])){
    activity_guessNA$steps[n]=step_interpolation(n);
  }
}
```
Histogram of the total number of steps taken each day

```{r}
sumofstepperday2<-ddply(activity_guessNA,c("date"),summarise,
                        totalsteps=sum(steps,na.rm = TRUE))
hist(sumofstepperday2$totalsteps,xlab="SumOfStepsPerDay",main="Histogram of the total number of steps taken each day")
```

```{r,echo=FALSE}
mean2<-mean(sumofstepperday2$totalsteps, na.rm=TRUE)
median2<-median(sumofstepperday2$totalsteps)
```
Mean `r mean2` and Median `r median2` of total number of steps per day after merging NAs.

Impact: Mean `r mean` and Median `r median` of total number of steps per day with NAs not merged.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
#head(activity_guessNA)

paindays= c("Monday","Tuesday","Wednesday","Thursday","Friday")

date_1<-ymd(activity_guessNA$date)

activity_guessNA$weekday<-as.factor(ifelse(weekdays(date_1)%in%paindays,"weekday","weekend"))

#head(activity_guessNA$weekday)

stepsperinterval.weekdaysplit<-ddply(activity_guessNA, c("interval","weekday"),summarise,
                    meansteps = mean(steps,na.rm=TRUE))

weekdayplot<-ggplot(stepsperinterval.weekdaysplit,aes(x=interval,y=meansteps))+
  facet_wrap(~weekday,nrow=2,ncol=1)+
  geom_line()+
  ggtitle("Mean steps over each 5min interval split by weekday/weekend")+
  ylab("Mean steps")+
  xlab("Interval number")
print(weekdayplot)

```


